---
title: "VIS Assignment 3 - Liu"
author: "Summer Liu"
date: "9/24/2020"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning = FALSE, results = "hide", message = FALSE)
```

# Data Load

```{r}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(maps)
library(dplyr)
library(knitr)
```


```{r}
BostonNeighborhood <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", 
                       quiet = TRUE) 

Trees <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/ce863d38db284efe83555caf8a832e2a_1.kml", 
                       quiet = TRUE)

PublicSchool <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", 
                       quiet = TRUE)

Hydrography <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2b3c0fa13b1c468eb702a3645fcc6bcb_5.kml", 
                       quiet = TRUE)

OpenSpace <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2868d370c55d4d458d4ae2224ef8cddd_7.kml",
                     quiet = TRUE)
```


```{r}
MA_State_Plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

BostonNeighborhood <- BostonNeighborhood %>%
  st_transform(MA_State_Plane)

Trees <- Trees %>%
  st_transform(MA_State_Plane)

PublicSchool <- PublicSchool %>%
  st_transform(MA_State_Plane)

Hydrography <- Hydrography %>%
  st_transform(MA_State_Plane)

OpenSpace <- OpenSpace %>%
  st_transform(MA_State_Plane)
```

## Overall map visualization

```{r}
ggplot(BostonNeighborhood) +
  geom_sf(fill = "lightblue", color = NA) + 
  geom_sf(data = Hydrography, color = "lightblue", size = 1) +
  geom_sf(data = PublicSchool, color = "red", size = 0.5) + 
  geom_sf(data = Trees, color = "springgreen4", size = 0.001)
  theme_map()
  annotation_scale()
```


```{r}
school_buffer <- st_buffer(PublicSchool, dist = 250) %>%
  st_union()

ggplot(school_buffer) + 
  geom_sf(color = "orange1") +
  theme_map()
```

```{r}
trees_school <- Trees[school_buffer,]

ggplot(school_buffer) + 
  geom_sf() +
  geom_sf(data = trees_school,
          color = "springgreen4",
          size = 0.01) +
  theme_map()
```

### Maps based on Carole's code

```{r}
Trees <- Trees %>%
  st_join(trees_school) %>%
    mutate(by_PublicSchool = !is.na(Name.y))

n_trees_school <- sum(Trees$by_PublicSchool)
n_trees_school
```

```{r}
n_trees <- length(Trees$by_PublicSchool)

pct_school_trees <- n_trees_school / n_trees

pct_school_trees
```
```{r}
left_side  <- st_bbox(Trees)$xmin
top_side <- st_bbox(Trees)$ymax

ggplot(Hydrography) +
  geom_sf(fill = "lightblue", color = NA) +
  geom_sf(data = Trees, size = 0.01,
          aes(color = by_PublicSchool)) +
  scale_color_manual(values = c("lightgreen", "darkgreen"),
          name = "Trees around Boston Public Schools",
          labels = c("Number of Trees within 250m",
                     "School within 250 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(panel.background = element_rect(fill = "cornsilk1"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

```{r}
BostonNeighborhood <- BostonNeighborhood %>%
  mutate(num_PublicSchool = lengths(st_covers(BostonNeighborhood, PublicSchool)))

ggplot(BostonNeighborhood) +
  geom_sf(color = NA, 
          aes(fill = num_PublicSchool)) +
  scale_fill_viridis_c(name = "Boston neighborhoods\nby number of public schools",
                       breaks = breaks <- seq(0, 80, by = 5),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Schools")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
```

### My own mappings

I would like to show the overlaying of Public schools and greenery/ Open spaces of conservation and recreation interests. 

```{r}
ggplot(BostonNeighborhood) +
  geom_sf(fill = "lightgray", color = NA) + 
  geom_sf(data = OpenSpace, color = "forestgreen", size = 1) +
  geom_sf(data = PublicSchool, color = "brown", size = 2) + 
  theme_map()
  annotation_scale()
```
I wanted to see the amount of parks/ open spaces per neighborhood, and this map is showing a highly identical pattern with the public school one. 

```{r}
BostonNeighborhood <- BostonNeighborhood %>%
  mutate(num_OpenSpace = lengths(st_covers(BostonNeighborhood, OpenSpace)))

ggplot(BostonNeighborhood) +
  geom_sf(color = NA, 
          aes(fill = num_OpenSpace)) +
  scale_fill_viridis_c( option = "plasma",
                        name = "Boston neighborhoods\nby number of open spaces",
                       breaks = breaks <- seq(0, 1000, by = 20),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "OpenSpaces")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
theme_map() +
  theme(legend.position = c(0.6,0))
guides(fill=guide_legend(title="Open Spaces around Boston"))
```

By this map, I am trying to show the number of trees per open space, in contrast to the last map, this one shows that though some areas have very few parks, but the quantity of trees is higher. 

```{r}
OpenSpace <- OpenSpace %>%
  mutate(num_Trees = lengths(st_covers(OpenSpace, Trees))) 
  
ggplot(OpenSpace) +
  geom_sf(color =NA,
          aes(fill = num_Trees)) +
  scale_fill_viridis_c(name = "Number of Trees in Open Spaces",
                       breaks = breaks <- seq(0, 5000, by = 500),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "trees")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.position = c(0.6,0))
 
```

Rather than quantity, in this one I calculated tree density per open space, and curious to find out why the areas showed with higher quantity don't also have higher density.

```{r}
OpenSpace <- OpenSpace %>%
  mutate(num_Trees = lengths(st_covers(OpenSpace, Trees))) %>%
  mutate(area = set_units(st_area(OpenSpace), km^2)) %>%
   mutate(Tree_dens = as.numeric(num_Trees / area))
 
ggplot(OpenSpace) +
  geom_sf(color = NA, 
          aes(fill = Tree_dens)) +
    scale_fill_viridis_c(option = "plasma",
                         name = "Tree Density per OpenSpace",
                       breaks = breaks <- seq(0, 80000, by = 10000),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Trees per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("Gray", 0.5), 
                                         color = "gray"))
```

This one I calculated public school density around different neighborhoods.

```{r}
BostonNeighborhood <- BostonNeighborhood %>%
  mutate(num_schools = lengths(st_covers(BostonNeighborhood, PublicSchool))) %>%
  mutate(area = set_units(st_area(BostonNeighborhood), km^2)) %>%
   mutate(School_dens = as.numeric(num_schools / area))
 
ggplot(BostonNeighborhood) +
  geom_sf(color = NA, 
          aes(fill = School_dens)) +
    scale_fill_viridis_c(name = "Public School Density per Neighborhood",
                       breaks = breaks <- seq(0, 20, by = 2),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "Schools per square km")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  theme_map() +
  theme(legend.position = "right",
    legend.background = element_rect(fill = alpha("Gray", 0.5), 
                                         color = "gray"))
 
```

In this one, I tried to calculate the average distance between public schools and openspaces. 

```{r}
PublicSchool <- PublicSchool %>%
  mutate(openspace_dist = st_nn(PublicSchool, OpenSpace, 
                           returnDist = TRUE)$dist) %>%
  mutate(openspace_dist = as.numeric(openspace_dist))

avg_tree_dist <- mean(PublicSchool$openspace_dist)

avg_tree_dist
```

```{r}
right_side <- st_bbox(PublicSchool)$xmax
left_side  <- st_bbox(PublicSchool)$xmin
top_side <- st_bbox(PublicSchool)$ymax
bottom_side <- st_bbox(PublicSchool)$ymin

ggplot(OpenSpace) +
  geom_sf(fill = "olivedrab4", color = NA) +
  geom_sf(data = PublicSchool, size = 2, color = "brown",
          aes(color = openspace_dist)) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c() +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 200, 
           y = top_side - 1800, 
           label = paste("Average distance between schools and\n openspaces is 117 meters."),
           hjust = 0, vjust = 0, size = 4) +
  theme_map() +
  theme(legend.position = "right") +
  theme(panel.background = element_rect(fill = "white"))
```













